---
title: "tidy-model"
author: "Xuan Bach Nguyen"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('tidyverse') ; library('modeldata')
```

## Introduction

This R markdown document demonstrates my abilities to use models for data analysis using a data set collected on crickets.

```{r load data, include=FALSE}
data(crickets, package = "modeldata")
names(crickets)
```
## What is that sound?

In this report, we examine what factors predict a cricket's chirp rate.

```{r summary, echo=FALSE}
summary(crickets)
ggplot(crickets, aes(x = rate)) +
      geom_histogram(bins = 8) +
      ggtitle("Distribution of the chirp rate of crickets") +
      xlab('Chirp rate(per min.)')
```
## Temperature affects chirp rate

```{r temp, echo=FALSE}
ggplot(crickets, aes(x = temp, y = rate)) +
      geom_point() +
      geom_smooth(method = 'lm') +
      ggtitle("Plot of tempaerature and chirp rate") +
      ylab('Chirp rate (per min.)') + 
      xlab('Temperature (Celcius)')

temp_lm <- lm(rate ~ temp, crickets)
summary.lm(temp_lm)
```

Based on a scatter plot of temperature and chirping and a correlation test, it seems that as temperature increases one degree, the rate of chirping increases about 4.2 chirps per minute.

## Species-specific effects of temperature on chirping

```{r species, echo=FALSE}
ggplot(crickets, aes(x = temp, y = rate, color = species)) +
      geom_point() +
      geom_smooth(method = 'lm') +
      ggtitle("Plot of temperature and chirp rate for two species of crickets") +     ylab('Chirp rate (per min.)') +
      xlab('Temperature (Celcius)')

species_lm <- lm(rate ~ temp + species, crickets)

summary.lm(species_lm)
```
      
O. exclamationis chirps ten times more frequently than O. niveus at the same temperature, according to the second model, which increases accuracy by taking species differences into consideration. It is more dependable than the general model since it forecasts chirp rates more accurately, as seen by its higher R-squared value (0.9896).

## Interaction

```{r species histogram, echo=FALSE}
ggplot(crickets, aes(x = rate, fill = species)) +
      geom_histogram(position = 'identity', alpha = 0.7, bins = 8) +
      ggtitle("Distribution of the chirp rate of crickets") + 
      xlab('Chirp rate (per min.)')
```

```{r interactions, echo=FALSE}
ggplot(crickets, aes(x = temp, y = rate, color = species)) +
      geom_point() +
      geom_smooth(method = 'lm') +
      ggtitle("Plot of temperature and chirp rate for two species of crickets") +
  
      ylab('Chirp rate (per min.)')
      xlab('Temperature (Celcius)')


species_x_temp_lm <- lm(rate ~ temp + species + temp:species, crickets)
summary.lm(species_x_temp_lm)
anova(species_lm, species_x_temp_lm)
```

## Predict diamond prices

```{r load diamonds, echo = FALSE}
data("diamonds", package = "ggplot2")
```
```{r diamond summary ,echo=FALSE}
summary('diamonds')
ggplot(diamonds, aes(x = price)) + 
      geom_histogram(alpha = 0.7) +
      ggtitle("Distribution of Diamond Prices") +
      xlab('Price (USD)')
```

## Relationship Between Carat and Price

```{r carrat and price, echo=FALSE}
ggplot(diamonds, aes(x = carat, y = price, color = 'cut')) + 
      geom_point(alpha = 0.3) + 
      geom_smooth(method = 'lm') + 
      ggtitle("Relationship Between Carat and Price by Cut") + 
      ylab('Price (USD)') +
      xlab('Carat')

diamond_lm <- lm(price ~ carat * cut + clarity, data = diamonds)
summary.lm(diamond_lm)
```
